---
title: "Statistical Inference Course Project Part 1"
author: "Keng Yew Hoe"
date: "10/13/2020"
output: pdf_document
---

## Overview
The purpose is to investigate the exponential distribution in R and comparing it with the Central Limit Theorem. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. with lambda = 0.2 for all the simulations. The investigation compares the distribution of averages of 40 exponentials over a thousand simulations

## Simulations
For the 1000 simulations, lambda is 0.2 and sample size is 40
```{r}
#Set seed to ensure reproducability 
set.seed(6666)

#Set lambda
lambda <- 0.2

#Set exponentials
n <- 40

#Simulating the exponential distribution and getting the mean of 1 simulation 
simul <- rexp(n,lambda)
simul
expMean <- mean(simul)
expMean

```
## Sample Mean versus Theoretical Mean
```{r}
#Getting the means of each simulations
simul1000 <- as.data.frame(replicate(1000,mean(rexp(n, lambda))))
names(simul1000) <- c("sample.mean")

#Calculating the mean of the means of 1000 simulation
mean1000 <- mean(simul1000$sample.mean)
mean1000

#The theoretical mean 
theoMean <- 1/lambda
theoMean
```
The theoretical mean and the experimental mean have a very similar value therefore the center of distribution of sample means of 40 exponential is close to the theoretical center of the distribution


## Sample Variance versus Theoretical Variance
```{r}
# Calculating the variance of this simulation
var1000 <- var(simul1000$sample.mean)
var1000

# The theoretical variance
theoVariance <- ((1/lambda)^2)/40
theoVariance
```
The variance of the distribution of means have a very similar value to the theoretical variance which is calculated by squaring the standard deviation and dividing by the sample size

## Distribution
```{r}
library(ggplot2)
ggplot(simul1000, aes(x=sample.mean)) +
    geom_histogram(aes(y = ..density..),colour="black",fill="cyan")+
    stat_function(fun=dnorm,args=list( mean=mean1000, sd=sqrt(var1000)),geom="line",color = "black", size = 1.0)+
    ggtitle("Histogram of the 1000 Simulation Samples Means") +
    scale_x_continuous("Simulation means")+
    ylab("Density")
```
The black curve represents the normal distribution which is used to compare with the histogram.
The central limit theorem states that the sample means would become that of a standard normal distribution as the sample size increases whilst meeting the two conditions of independence ( n < 10% ) and normal, or if skewed distribution, that n > 30.

* * *

