---
title: "Statistical Inference Course Project Part 2"
author: "Keng Yew Hoe"
date: "10/13/2020"
output: pdf_document
---
## Overview
Analyzing the ToothGrowth data in the R datasets package while providing a basic summary of the data and using condifence intervals with hypothesis tests to compare tooth growth by supp and dose.

##1. Load the ToothGrowth data and perform some basic exploratory data analyses
```{r}
# Load the ToothGrowth data
data(ToothGrowth)
str(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)

# Perform Exploratory Data Analysis
library(ggplot2)

#Replacing OJ and VC with Orange Juice and Vitamin C
ToothGrowth$supp <- as.character(ToothGrowth$supp)
ToothGrowth$supp[ToothGrowth$supp == "OJ"] <- "Orange Juice"
ToothGrowth$supp[ToothGrowth$supp == "VC"] <- "Vitamin C"

#Plotting
ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  facet_grid(.~supp) +
  geom_boxplot(aes(fill = supp), show_guide = FALSE) +
  labs(title="Tooth length along dosage for each type of supplement", 
    x="Dose (mg/day)",
    y="Tooth Length")
g <- ggplot(aes(x = " ", y = len), data = ToothGrowth)
g + geom_boxplot(aes(fill=supp))+ xlab(" ") +
        ylab("Tooth Length") + facet_grid(~ dose) + ggtitle("Tooth Length Along Supplement Type")

```

## 2. Basic summary of the data
Based on the graph, tooth growth increases as dosage increases.It is also shown that orange juice has a higher median tooth length when dosage is 0.5 and 1 mg/day. Both types of supplements have the same median when dosage is 2 mg/day.

## 3. Use confidence intervals & hypothesis tests to compare tooth growth by supplement and dose
### Hypothesis 1
Orange juice & Vitamin C deliver the same tooth growth across the data set.
```{r}
h1<-t.test(len ~ supp, ToothGrowth)
h1$conf.int
h1$p.value

```
The confidence intervals includes 0 and the p-value is greater than the threshold of 0.05.  The null hypothesis cannot be rejected.

### Hypothesis 2
For the dosage of 0.5 mg/day, the two supplements deliver the same tooth growth.
```{r}
h2<-t.test(len ~ supp, subset(ToothGrowth, dose == 0.5))
h2$conf.int
h2$p.value
```
The confidence interval does not include 0 and the p-value is below the 0.05 threshold. The null hypothesis can be rejected. The alternative hypothesis that 0.5 mg/day dosage of orange juice delivers more tooth growth than Vitamin C is accepted.

### Hypothesis 3
For the dosage of 1 mg/day, the two supplements deliver the same tooth growth
```{r}
h3<-t.test(len ~ supp, data = subset(ToothGrowth, dose == 1))
h3$conf.int
h3$p.value
```
The confidence interval does not include 0 and the p-value is smaller than the 0.05 threshold. The null hypothesis can be rejected. The alternative hypothesis that 1 mg/day dosage of orange juice delivers more tooth growth than Vitamin C is accepted.

### Hypothesis 4
For the dosage of 2 mg/day, the two supplements deliver the same tooth growth
```{r}
h4<-t.test(len ~ supp, data = subset(ToothGrowth, dose == 2))
h4$conf.int
h4$p.value
```
The confidence interval does include 0 and the p-value is larger than the 0.05 threshold. The null hypothesis cannot be rejected.


## State your conclusions and the assumptions needed for your conclusions
Orange juice delivers more tooth growth than Vitamin C for dosages 0.5 & 1.0 mg/day. Orange juice and Vitamin C deliver the same amount of tooth growth for dose amount 2.0 mg/day.
For the entire data set we cannot conclude orange juice is more effective than vitamin C however it can be concluded that dosage will cause an increase in tooth length

Assumptions

* Normal distribution of the tooth lengths
* No other unmeasured factors are affecting tooth length